## Exploring the relationship between type of transmission and miles per gallon

author: *Grigory Miholap*  
date: *20/11/2015*

## 1. Overview
There is relationship between type of transmission and miles per gallon. Average value `mpg` for manual transmission is 24 miles per gallon, and for automatic transmission - 17 miles per gallon. Difference between groups is significant and equal 7. And can say, that manual transmission is better for MPG, than automatic transmission and average difference equal 7 miles per gallon. But should note, that our model don't satisfy the assumptions "Homoskedasticity" and "Independence of errors".

## 2. Loading the data
```{r echo=FALSE, results='hide'}
library(ggplot2)
library("xtable")
library("car")
```
```{r}
data(mtcars)
```


**Note 1:** I've hidden some code chunks to make report shorter. All code available [here](https://github.com/Grag2015/Coursera/blob/master/6_Regression%20Models/project.R)

```{r echo=FALSE, results='hide'}
# my functions
printlm <- function(model){
  tempsum <- summary(model)
  cat("Residual standard error:", format(signif(tempsum$sigma, 
                                                4)), "on", tempsum$df[2L], "degrees of freedom")
  cat("\n")
  cat("Multiple R-squared:  ", round(tempsum$r.squared, digits=4),	"Adjusted R-squared:  ",
      round(tempsum$adj.r.squared, digits=4))
  cat("\n")
  cat("F-statistic: ", round(tempsum$fstatistic[1],2), "on", round(tempsum$fstatistic[2],0), 
      "and", round(tempsum$fstatistic[3],0), "DF,  p-value:",
      format.pval(pf(tempsum$fstatistic[1L],tempsum$fstatistic[2L], 
                     tempsum$fstatistic[3L], lower.tail = FALSE)))
}
# пользовательские функции для график попарных корреляций
    panel.density <- function(x, ...) {
        n.groups <-  1
        adjust <-  1
        groups = NULL
        if (n.groups > 1) {
            levs <- levels(groups)
            for (i in 1:n.groups) {
                xx <- x[levs[i] == groups]
                dens.x <- try(density(xx, adjust = adjust, na.rm = TRUE), 
                  silent = TRUE)
                if (!inherits(dens.x, "try-error")) {
                  lines(dens.x$x, min(x, na.rm = TRUE) + dens.x$y * 
                    diff(range(x, na.rm = TRUE))/diff(range(dens.x$y, 
                    na.rm = TRUE)), col = col[i])
                }
                else warning("cannot estimate density for group ", 
                  levs[i], "\n", dens.x, "\n")
                rug(xx, col = col[i])
            }
        }
        else {
            dens.x <- density(x, adjust = adjust, na.rm = TRUE)
            lines(dens.x$x, min(x, na.rm = TRUE) + dens.x$y * 
                diff(range(x, na.rm = TRUE))/diff(range(dens.x$y, 
                na.rm = TRUE)))
            rug(x)
        }
#         if (do.legend) 
#             legendPlot(position = if (is.null(legend.pos)) 
#                 "topright"
#             else legend.pos)
#         do.legend <<- FALSE
    }
panel.cor <- function(x, y, digits=2, prefix="", cex.cor, ...)
{
    usr <- par("usr"); on.exit(par(usr))
    par(usr = c(0, 1, 0, 1))
    r <- cor(x[!is.na(x*y)], y[!is.na(x*y)])
    txt <- format(c(r, 0.123456789), digits=digits)[1]
    txt <- paste(prefix, txt, sep="")
    if(missing(cex.cor)) cex.cor <- 0.8/strwidth(txt)
    text(0.5, 0.5, txt, cex = cex.cor * max(abs(r), 0.25))
}

panel.smooth2 <- function(x, y){
panel.smooth(x, y,iter = 1)
}
```

## 3. Analysis
In first, I'll see how variables `am` and `mpg` look
```{r}
head(mtcars[,c("am","mpg")])
```

```{r}
table(mtcars$am)
```

We see, that `am` is factor with 2 levels - (0 = automatic, 1 = manual)

```{r}
summary(mtcars$mpg)
```

and `mpg` is countable variable in interval [10.40-33.90]. So, we can use simple linear regression.

I'll show, that values `mpg` depend on type of transmission.
Next, I'm going to run t.test to check out significance of differences
```{r}
t.test(x=mtcars[mtcars$am==1,"mpg"], y=mtcars[mtcars$am==0,"mpg"], var.equal = F)
```

So, test shows that `p-value = 0.001374 < 0.05`, so we have significant differences between groups "automatic" and "manual"

I'll make the boxplot to show more details 
```{r}
ggplot(data=mtcars, aes(x=factor(am), y=mpg))+geom_boxplot()
```

Next, I'll fit a model with mpg as the outcome that includes type of transmission as a factor variable (predictor)
```{r}
fit <- lm(data = mtcars, mpg ~ factor(am))
```
```{r echo=FALSE, results='hold'}
printlm(fit)
```
```{r results='asis',  echo=FALSE}
xt <- xtable(summary(fit))
print(xt, type="html")
```

So we can write formula `mpg = 17.147 + am*7.24`, where `am` equals either `1` or `0`. If `am = 0` (automatic transmission), then `mpg = 17.147 miles/gallon`, if `am = 1` (manual transmission), then `mpg = 17.147 + 7.24 = 24.387 miles/gallon`. We see, that coefficients are significant, because `p-value << 0.05` for both coefficients. 

## 4. Diagnostic
**1.** normality of the errors
```{r}
qqPlot(fit, labels=row.names(mtcars), id.method="identify",
       simulate=TRUE, main="Q-Q Plot")
```

QQ-plot let us investigate normality of the errors. 
Distribution of the errors likes Gaussian, Shapiro-Wilk normality test can get precise answer
```{r}
shapiro.test(fit$residuals)
```

Value `p-value = 0.8573 >> 0.05` means, that the errors have normal distribution. Also we can estimate mean and standart deviation of this distribution, 
`mean=mean(fit$residuals)=``r mean(fit$residuals)`~0 and `sd=sd(fit$residuals)=``r sd(fit$residuals)`

**2.** Homoskedasticity
```{r results='hide', warning=FALSE, message=FALSE}
spreadLevelPlot(fit)
```

We see, that variance doesn't appear to be constant.  

**3.** Independence of the errors
```{r}
cor(fit$residuals, mtcars$mpg)
```
So, we see strong correlation between the residuals and the predictors and saw.

**Note 2** We see, that our model don't satisfy the assumptions "Homoskedasticity" and "Independence of errors". I've tried take logs of variable `mpg`, but this transformation doesn't help. 

## 5. Summary
Now we can answer the questions of interest 

>Is an automatic or manual transmission better for MPG

manual transmission is better for MPG, than automatic transmission

>Quantify the MPG difference between automatic and manual transmissions

average difference equal 7 miles per gallon

**Final formula**

`mpg = 17.147 + am*7.24 + e`, where `am` equals either `1`(manual) or `0` (automatic), and
`e` from `N(0, 4.8^2)` (But don't forget Note 2)

______