---
title: "Assessment_2"
author: "Grag"
date: "15 ноября 2015 г."
output: html_document
---

setwd("d:/Grag/R/R-studio/Coursera/4_Reproducible Research/RepData_PeerAssessment2/")

## Title (that briefly summarizes your data analysis)

## Synopsis 
Immediately after the title, there should be a synopsis which describes and summarizes your analysis in at most 10 complete sentences.

## Data Processing
1. Данные скачал по ссылке [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) в 
архиве bz2.
2. Распаковал архив в рабочую директорию 
3. Read csv-file
```{r cache=TRUE}
df <- read.csv("./repdata-data-StormData.csv")
head(df)
summary(df$FATALITIES)
summary(df$INJURIES)
tbl <- table(df$EVTYPE)
tblord <- tbl[order(tbl, decreasing = T)]
```


## Results
рассчитаем суммарную, среднюю  и количество event для каждого типа события
```{r}
library(dplyr)
tmp <- group_by(df, EVTYPE)
tmp2 <- summarise(.data=tmp, FATALITIEStotal=sum(FATALITIES), FATALITIESmean=mean(FATALITIES),
                  EVTYPEn=n(), INJURIEStot=sum(INJURIES), 
                  INJURIESaver=mean(INJURIES))
# order by 
tmp2ord <- tmp2[order(tmp2$FATALITIEStotal, decreasing = T),]
# check NAs
sapply(tmp2ord, function(e) sum(is.na(e)))

# create new factor for FATALITIEStotal
tmp2ord$TotalFATALITIES <- character(length = nrow(tmp2ord))
tmp2ord[tmp2ord$FATALITIEStotal>=500,]$TotalFATALITIES<-"500+" 
tmp2ord[tmp2ord$FATALITIEStotal>=100 & tmp2ord$FATALITIEStotal<500,]$TotalFATALITIES<-"100-500" 
tmp2ord[tmp2ord$FATALITIEStotal>=50 & tmp2ord$FATALITIEStotal<100,]$TotalFATALITIES<-"50-100" 
tmp2ord[tmp2ord$FATALITIEStotal<50,]$TotalFATALITIES<-"0-50" 
tmp2ord$TotalFATALITIES <- factor(x=tmp2ord$TotalFATALITIES,levels = c("0-50", "50-100",
                                                                       "100-500", "500+"),
                                ordered = T)
# содержит записи с FATALITIEStotal>=50 - для дальнейшего использования на графике
tmp3 <- tmp2ord[tmp2ord$TotalFATALITIES!="0-50" & tmp2ord$TotalFATALITIES!="50-100",]
tmp3$EVTYPE <- as.character(tmp3$EVTYPE)
tmp3[tmp3$TotalFATALITIES=="500+",]$EVTYPE <- paste0(tmp3[tmp3$TotalFATALITIES=="500+",]$EVTYPE,
                                                     " (", tmp3[tmp3$TotalFATALITIES=="500+",]$FATALITIEStotal, ")")
```

как видим существует связь между частотой события и средним ущербом
```{r}
library(ggplot2)

ggplot(data = tmp2ord, aes(x=(FATALITIESmean)^(1/4), y=(EVTYPEn)^(1/4), col=TotalFATALITIES, size=(FATALITIEStotal)^(1/4))) + geom_point()+xlim(0,1.5)+ylim(0,24)+annotate("text", label = tmp3$EVTYPE, x = (tmp3$FATALITIESmean)^(1/4), y = (tmp3$EVTYPEn)^(1/4), size = 3, colour = "#2e3a23",  hjust=-0.1)+xlab("Mean fatalities per event, pcs^(1/4)")+ylab("Number of events, pcs^(1/4)")+
    labs(title="Relation between 'Mean fatalities per event' and 'Number of events'")+scale_size_continuous(name = "bigger size - bigger", labels = c("","","","",""), breaks=NULL)

ggplot(data = tmp2ord[1:150,], aes(x=log(FATALITIESmean+1), y=log(EVTYPEn), size=FATALITIEStotal)) + geom_point() + geom_smooth()

ggplot(data = tmp2ord[1:150,], aes(x=log(INJURIESaver+1), y=log(EVTYPEn), size=(INJURIEStot)^(1/4))) + geom_point()+ geom_smooth()

```

devtools::install_github("hadley/scales") 
